---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, echo=TRUE}
library(knitr)
opts_chunk$set(echo=TRUE)
```

## Loading and preprocessing the data

```{r}
library(readr)
library(lubridate)

df = read.csv(unzip('activity.zip','activity.csv'), 
   na.string='NA')

df$fint  = sprintf('%04i', df$interval)
df$dmin  = floor(df$int / 100) * 60 + (df$int %% 100)
df$rdate = ymd(df$date)

```

## What is mean total number of steps taken per day?

```{r}

df.narm = df[!is.na(df$steps),]
tstep = tapply(df.narm$steps, df.narm$date, sum)

stmean = format(mean(tstep), nsmall=2, digits=7, big.mark=',')
stmed  = format(median(tstep), big.mark=',')

hist(tstep,
   main='Histogram of Total Steps per Day', 
   xlab='Total Steps per Day', 
   ylab='Frequency') 

```

#### The mean of the total number of steps taken per day is `r stmean`.  
#### The median of the total number of steps taken per day is `r stmed`.      


## What is the average daily activity pattern?

```{r}
library(dplyr)

mstep = df.narm %>% group_by( dmin ) %>%
   summarize(mstep = mean(steps)) %>%
   arrange(dmin) %>% data.frame()

maxstep = format(max(mstep$mstep), nsmall=3, digits=5)
maxdmin = mstep$dmin[which.max(mstep$mstep)]

dmin1 = sprintf('%02i:%02i', 
   floor(maxdmin / 60), maxdmin %% 60)
dmin2 = sprintf('%02i:%02i', 
   floor((maxdmin+5) / 60), (maxdmin+5) %% 60)

plot(mstep$dmin,mstep$mstep, type='l', axes=F, 
   main='Average Steps in 5-Min Interval in Each Day',
   xlab='Time of Day (HH:MM in the 24-hour format)',
   ylab='Steps in 5-Min Interval')
axis(1, at=0:24*60, labels=F, tck=-0.01)
axis(1, at=0:8*180, labels=sprintf('%02i:00',0:8*3))
axis(2)

```

#### The maximum steps among the average steps taken in 5-min interval is: 
#### `r maxstep` steps during the period between `r dmin1` and `r dmin2`. 

## Imputing missing values

#### The number of rows with missing values is `r length(which(is.na(df$steps)))`.
#### Missing Value Mitigation Method  
* The missing values in weekday records are filled with the average value of 
the time period in other weekday records.  
* The missing values in weekend records are filled with the average value of 
the time period in other weekend records.  

```{r}

naidx = which(is.na(df$steps)) 
df$ww = ifelse(
   weekdays(as.Date(df$date)) %in% c('Saturday','Sunday'),
   'Weekend','Weekday' ) 
df$wwint = paste(df$ww, df$fint, sep='-')

avgstep = tapply(df$steps, df$wwint, mean, na.rm=T)

df.fill = df
df.fill$steps[naidx] = avgstep[ df$wwint[naidx] ]

tstep = tapply(df.fill$steps, df.fill$date, sum)

stmean2 = format(mean(tstep), nsmall=2, digits=7, big.mark=',')
stmed2  = format(median(tstep), big.mark=',')

hist(tstep,
   main='Histogram of Total Steps per Day', 
   xlab='Total Steps per Day', 
   ylab='Frequency') 

```

### BEFORE missing values are filled: 
#### The mean of the total number of steps taken per day is `r stmean`.  
#### The median of the total number of steps taken per day is `r stmed`.      

### AFTER missing values are filled: 
#### The mean of the total number of steps taken per day is `r stmean2`.  
#### The median of the total number of steps taken per day is `r stmed2`.      

## Are there differences in activity patterns between weekdays and weekends?

```{r fig.height=8}

mstep = df.fill %>% group_by( dmin, ww ) %>%
   summarize(mstep = mean(steps)) %>%
   arrange(dmin, ww) %>% data.frame()

mstep = reshape(mstep, direction='wide',
   idvar='dmin', timevar='ww')
names(mstep) = sub('mstep.','', names(mstep))

par(mfrow=c(2,1))

plot(mstep$dmin,mstep$Weekday, type='l', axes=F, 
   main='Average Steps in 5-Min Interval in WEEKDAY',
   xlab='Time of Day (HH:MM in the 24-hour format)',
   ylab='Steps in 5-Min Interval', ylim=c(0,250))
axis(1, at=0:24*60, labels=F, tck=-0.01)
axis(1, at=0:8*180, labels=sprintf('%02i:00',0:8*3))
axis(2)

plot(mstep$dmin,mstep$Weekend, type='l', axes=F, 
   main='Average Steps in 5-Min Interval in WEEKEND',
   xlab='Time of Day (HH:MM in the 24-hour format)',
   ylab='Steps in 5-Min Interval', ylim=c(0,250))
axis(1, at=0:24*60, labels=F, tck=-0.01)
axis(1, at=0:8*180, labels=sprintf('%02i:00',0:8*3))
axis(2)

```


